Tactics and Marks in Banach Mazur Games

### 2017-12-07 postdoctoral/k-markov-tactic-banach-mazur.tex
write up notes on marks/tactics

      \theoremstyle{definition}
      \newtheorem{definition}[theorem]{Definition}
      \newtheorem{example}[theorem]{Example}
      \newtheorem{game}[theorem]{Game}

@@

      \theoremstyle{definition}
      \newtheorem*{definition*}{Definition}
      \newtheorem*{example*}{Example}
      \newtheorem*{game*}{Game}

@@

\title{\(k\)-Limited Strategies in Banach Mazur Games}
\author{Steven Clontz}

@@

@@

  \begin{definition}
    \(\bmPoGame{\mb P}{W}\) is a game defined for all non-empty partial orders
    \(\mb P\) and all subsets \(W\subseteq\mb P^\downarrow\) closed under
    zipping. During round \(0\), \(\plI\) chooses \(a_0\in\mb P\),
    and then \(\plII\) chooses \(b_0\leq a_0\); during around \(n+1\),
    \(\plI\) chooses \(a_{n+1}\leq b_n\), and then \(\plII\) chooses
    \(b_{n+1}\leq a_{n+1}\). \(\plII\) wins this game if
    \(\<a_0,a_1,\dots\>\in W\).
  \end{definition}

  \begin{theorem}
    \(\plII\kmarkwin{(k+1)}\bmPoGame{\mb P}{W}\) if and only if
    \(\plII\ktactwin{(k+1)}\bmPoGame{\mb P}{W}\).
  \end{theorem}

  \begin{proof}
    Let \(\tau(\vec p,n+1)\) be a winning \((k+1)\)-mark for \(\plII\),
    where \(n+1\) is the number of moves previously made by \(\plI\).
    Let \(\preceq\) well-order \(\mb P^{k+1}\).
    For \(f\in\mb P^\omega\), let
    \(f_n=\<f(n),\dots,f(n+k)\>\in\mb P^{k+1}\).

    For \(\vec p\in \mb P^{k+1}\) and \(q\in\mb P\), say \(\vec p\) is
    \(n\)-above \(q\) if there exists \(s_n(\vec p)\in\mb P\)
    such that
    \[
      \vec p(k)
    \geq
      s_n(\vec p)
    \geq
      \tau(\vec r\concat\<s_n(\vec p)\>,n+k+1)
    \geq
      q
    \]
    for all \(\vec r\in\mb P^k\) ``between \(\vec p(0)\) and \(\vec p(k)\)''.

    For \(\vec p\in\mb P^{k+1}\), let \(t^0(\vec p)=\vec p\),
    \(\tau^{n+1}(\vec p)=\tau(t^n(\vec p),n+1)\), and
    \(t^{n+1}(\vec p)=(t^n(\vec p)\last k)\concat\<\tau^{n+1}(\vec p)\>\).

    Say \(\vec p\) is \(\omega\)-above \(q\) if \(\vec p\) is \(n\)-above
    \(q\) for all \(n<\omega\). If \(\vec p\) is \(\omega\)-above some
    \(l(\vec p)\), then say \(\vec p\) is long; otherwise call \(\vec p\)
    short.

    For \(\vec p\) short, let
    \[
      \mu(\vec p)
        =
      \min_{\preceq}\{
        \vec r\text{ short}
          :
        \vec{r}(k)\geq\vec{p}(k)
      \}
    \]
    and since \(\mu(\vec p)\) is not \(n\)-above \(\vec p(k)\) for some \(n\),
    let
    \[
      N(\vec p)
        =
      \min\{
        n<\omega
      :
        \mu(\vec p)\text{ is not }n\text{-above }\vec p(k)
      \}
    .\]

    We define
    \[
      \sigma(\vec p)
        =
      \begin{cases}
        \tau(\vec p,|\vec p|) & |\vec p|\leq k \\
        l(\vec p) & \vec p \text{ is long} \\
        \tau^{N(\vec p)+1}(\vec p) & \vec p \text{ is short}
     \end{cases}
    .\]

    Suppose \(\sigma\) is legally attacked by \(a\in\mb P^\omega\).
    % Suppose \(b\) is a subsequence of \(a\) where
    % \(b_n\) is long for all \(n<\omega\).
    % Since \(b_n\) improves \(a_n\)
    % for all \(n<\omega\), it follows that
    % \(a_n\) is long for all \(n<\omega\).
    Thus for \(n<k\),
    \[
      a(n)
        \geq
      \tau(a\rest(n+1),n)
        =
      \sigma(a\rest(n+1))
        \geq
      a(n+1)
    .\]
    For \(n\leq\omega\), if \(a_n\) is long, then \(a_n\) is \(n\)-above
    \(l(a_n)\). Therefore,
    \[
      a(n+k)
        =
      a_n(k)
        \geq
      s_n(a_n)
        \geq
      \tau((a_n\last k)\concat\<s_n(a_n)\>,n+k)
        \geq
      l(a_n)
        =
      \sigma(a_n)
        \geq
      a(n+k+1)
    .\]
    Thus if \(a_n\) is long for \(n<\omega\), it follows that  \(\<a(0),\dots,a(k-1),s_0(a_0)(k),s_1(a_1)(k),\dots\>\)
    is a legal attack against \(\tau\). Since \(\tau\) is winning,
    this attack belongs to \(W\). Since this attack zips into \(a\),
    \(a\) also belongs to \(W\).

    Otherwise, we may choose \(k<N<\omega\) such that
      \begin{itemize}
        \item \(a_{n+N}\) is short for all \(n<\omega\),
              since \(a_m\) short implies \(a_n\) short for all \(m\leq n\).
        \item \(\mu(a_{n+N})=\vec m\) is fixed for all \(n<\omega\), since
              there cannot be an infinite \(\preceq\)-decreasing sequence.
      \end{itemize}
    As a result, \(a_{n+N}\) is...
    Thus for \(n<k\),
    \[
      b(n)
        \geq
      \tau(b\rest(n+1),n)
        =
      \sigma(a\rest(n+1))
        \geq
      a(n+1)
    .\]

  \end{proof}

>>>>

      \theoremstyle{definition}
      \newtheorem{definition}[theorem]{Definition}
      \newtheorem{example}[theorem]{Example}
      \newtheorem{observation}[theorem]{Observation}
      \newtheorem{game}[theorem]{Game}

@@

      \theoremstyle{definition}
      \newtheorem*{definition*}{Definition}
      \newtheorem*{example*}{Example}
      \newtheorem*{observation*}{Observation}
      \newtheorem*{game*}{Game}

@@

\title{Tactics and Marks in Banach Mazur Games}
\author{Steven Clontz}

@@

  My notes on Galvin/Telgarsky's Theorem 5 from \cite{MR831181}.

@@

  \begin{definition}
    \(\bmPoGame{\mb P}{W}\) is a game defined for all non-empty partial orders
    \(\mb P\) and all subsets \(W\subseteq\mb P^\downarrow\).
    During round \(0\), \(\plI\) chooses \(a_0\in\mb P\),
    and then \(\plII\) chooses \(b_0\leq a_0\); during around \(n+1\),
    \(\plI\) chooses \(a_{n+1}\leq b_n\), and then \(\plII\) chooses
    \(b_{n+1}\leq a_{n+1}\). \(\plII\) wins this game if
    \(\<a_0,a_1,\dots\>\in W\).
  \end{definition}

  \begin{theorem}
    Let \(W\subseteq\mb P^\downarrow\) be closed under zipping.
    \(\plII\markwin\bmPoGame{\mb P}{W}\) if and only if
    \(\plII\tactwin\bmPoGame{\mb P}{W}\).
  \end{theorem}

  \begin{proof}
    Let \(\tau(p,n+1)\) be a winning mark for \(\plII\), where \(p\)
    is the most recent move by \(\plI\) and \(n+1\)
    is the number of moves made by \(\plI\).
    Define \(\tau^0(p)=p\) and \(\tau^{n+1}(p)=\tau(\tau^n(p),n+1)\).
    Let \(\preceq\) well-order \(\mb P\).

    For \(p,q\in \mb P\), say \(p\geq_n q\) if there exist
    \(s_m(p)\in\mb P\) for \(m\leq n\)
    such that
    \[
      p
    \geq
      s_m(p)
    \geq
      \tau(s_m(p),n+1)
    \geq
      q
    .\]
    Note that \(p'\geq p\geq_n q\geq q'\) implies \(p'\geq_n q'\),
    and \(p\geq_n \tau^n(p)\).

    Say \(p\geq_\omega q\) whenever \(p\geq_n q\) for all \(n<\omega\).
    If \(p\geq_\omega l(p)\) for some \(l(p)\), then say \(p\) is long;
    otherwise call \(p\) short.

    For \(p\) short, let
    \[
      \mu(p)
        =
      \min_{\preceq}\{
        r\text{ short}
          :
        r\geq p
      \}
    \]
    and since \(\mu(p)\not\geq_n p\) for some \(n\), let
    \[
      N(p)
        =
      \min\{
        n<\omega
      :
        \mu(p)\not\geq_n p
      \}
    .\]
    Note that whenever \(\mu(p)=\mu(q)\) for \(p\geq_n q\),
    it follows that \(\mu(p)\geq_n q\) and therefore \(N(p)<N(q)\).

    We define
    \[
      \sigma( p)
        =
      \begin{cases}
        l(p) & p \text{ is long} \\
        \tau^{N(p)+1}( p) &  p \text{ is short}
     \end{cases}
    .\]

    Suppose \(\sigma\) is legally attacked by \(a\in\mb P^\omega\).
    For \(n\leq\omega\), if \(a(n)\) is long, then \(a(n)\geq_n l(a(n))\).
    Therefore,
    \[
      a(n)
        \geq
      s_n(a(n))
        \geq
      \tau(s_n(a(n)),n+1)
        \geq
      l(a(n))
        =
      \sigma(a(n))
        \geq
      a(n+1)
    .\]
    Thus if \(a(n)\) is long for \(n<\omega\), it follows that
    \(c\in\mb P^\downarrow\) defined by \(c(n)=s_n(a(n))\)
    is a legal attack against \(\tau\). Since \(\tau\) is winning,
    \(c\in W\), and since \(c\) zips into \(a\),
    \(a\in W\) as well.

    Otherwise, we may choose a final subsequence \(b\) of \(a\) such that
      \begin{itemize}
        \item \(b(n)\) is short for all \(n<\omega\),
              since \(a(m)\) short implies \(a(n+m)\) short for all
              \(n<\omega\).
        \item \(\mu(b(n))=\mu'\) is fixed for all \(n<\omega\), since
              there cannot be an infinite \(\preceq\)-decreasing sequence.
      \end{itemize}
    As a result,
    \[
      b(n)
        \geq_{N(b(n))}
      \tau^{N(b(n))+1}(b(n))
        =
      \sigma(b(n))
        \geq
      b(n+1)
    \]
    and therefore \(N(b(n))<N(b(n+1))\). In particular, \(N(b(n))\geq n\).

    Thus for \(n<\omega\),
    \[
      b(n)
        \geq
      \tau^{n}(b(n))
        \geq
      \tau(\tau^{n}(b(n)),n+1)
        \geq
      \tau^{N(b(n))+1}(b(n))
        =
      \sigma(b(n))
        \geq
      b(n+1)
    .\]
    As a result, \(c\in\mb P^\downarrow\) defined by \(c(n)=\tau^n(b(n))\)
    is a legal attack against the winning strategy \(\tau\). Therefore
    \(c\in W\), and since \(c\) zips into \(b\) and \(a\), we conclude
    \(a\in W\).
  \end{proof}

  \begin{observation}
    When \(\mb P=T(X)\setminus\{\emptyset\}\) is ordered by set-inclusion
    and \(W=\{U\in\mb P^\downarrow:\bigcap_{n<\omega}U(n)\not=\emptyset\}\),
    then \(\bmPoGame{\mb P}{W}\) is exactly the topological Banach Mazur game
    \(\bmGame{X}\). Note \(W\) is closed under zipping.
  \end{observation}

  \begin{corollary}
  \(\plII\markwin\bmGame{X}\) if and only if
  \(\plII\tactwin\bmGame{X}\).
  \end{corollary}

  \begin{observation}
    When \(\mb P=\{(U,x):U\in T(X)\setminus\{\emptyset\},x\in U\}\)
    is ordered by \((U,x)\geq(V,y)\) whenever \(x\in V\subseteq U\)
    and \(W=\{\<U(n),x(n)\>_{n<\omega}\in\mb
    P^\downarrow:\bigcap_{n<\omega}U(n)\not=\emptyset\}\),
    then \(\bmPoGame{\mb P}{W}\) is almost the Choquet game, except the first
    player also gets to choose a point. Note \(W\) is closed under zipping.
  \end{observation}

  \bibliographystyle{plain}
  \bibliography{../bibliography}
